---
title: "EDA: THEW HR Time Series"
subtitle: "Exploratory data analysis of 1-min and 5-min downsampled heart rate (CAD, Healthy, ESRD)"
date: today
format:
  gfm:
    toc: true
    toc-depth: 3
  html:
    self-contained: true
    toc: true
    toc-depth: 3
    code-fold: true
execute:
  warning: false
  message: false
  echo: false
---

## 1 Setup & Data Loading

```{r setup}
library(tidyverse)
library(here)
library(knitr)
library(scales)

theme_set(theme_minimal(base_size = 12))
```

```{r load-data}
thew_1min <- read_csv(here("data_processed/thew/hr_timeseries_1min/panel_24h_1min.csv"),
                      show_col_types = FALSE)
thew_5min <- read_csv(here("data_processed/thew/hr_timeseries_1min/panel_24h_wearable5.csv"),
                      show_col_types = FALSE)
```

```{r load-summary}
summary_file <- here("data_processed/thew/hr_timeseries_1min/combined_processing_summary.csv")
if (file.exists(summary_file)) {
  proc_summary <- read_csv(summary_file, show_col_types = FALSE)
  cat("Processing summary loaded:", nrow(proc_summary), "records\n")
} else {
  proc_summary <- NULL
  cat("Processing summary file not found — skipping.\n")
}
```

```{r data-overview}
cat("--- 1-minute data ---\n")
cat("Dimensions:", nrow(thew_1min), "rows x", ncol(thew_1min), "cols\n")
cat("Subjects:", n_distinct(thew_1min$id), "\n")
cat("Conditions:", paste(unique(thew_1min$condition), collapse = ", "), "\n\n")

cat("--- 5-minute data ---\n")
cat("Dimensions:", nrow(thew_5min), "rows x", ncol(thew_5min), "cols\n")
cat("Subjects:", n_distinct(thew_5min$id), "\n")
```

```{r glimpse}
glimpse(thew_1min)
```

## 2 Subject-Level Summary

### Counts by Condition and Dataset

```{r condition-counts}
thew_1min %>%
  distinct(id, condition, dataset) %>%
  count(condition, dataset, name = "n_subjects") %>%
  kable(caption = "Subjects by condition and source dataset")
```

```{r condition-totals}
thew_1min %>%
  distinct(id, condition) %>%
  count(condition, name = "n_subjects") %>%
  kable(caption = "Total subjects by condition")
```

### Processing Summary

```{r proc-summary}
if (!is.null(proc_summary)) {
  proc_summary %>%
    group_by(condition) %>%
    summarise(
      n_subjects       = n(),
      mean_duration_hr = mean(duration_hours, na.rm = TRUE),
      mean_valid_min   = mean(n_valid_minutes, na.rm = TRUE),
      mean_hr_overall  = mean(mean_hr, na.rm = TRUE),
      n_success        = sum(status == "success", na.rm = TRUE),
      .groups = "drop"
    ) %>%
    kable(digits = 1, caption = "Processing summary by condition")
}
```

### Recording Duration

```{r recording-duration}
dur <- thew_1min %>%
  group_by(id, condition, dataset) %>%
  summarise(
    n_minutes   = n(),
    n_valid_min = sum(valid == TRUE, na.rm = TRUE),
    duration_hr = n_minutes / 60,
    .groups = "drop"
  )

dur %>%
  group_by(condition) %>%
  summarise(
    n_subjects       = n(),
    median_minutes   = median(n_minutes),
    min_minutes      = min(n_minutes),
    max_minutes      = max(n_minutes),
    median_valid_min = median(n_valid_min),
    .groups = "drop"
  ) %>%
  kable(caption = "Recording duration by condition")
```

## 3 Missing Data Audit

### NA and Invalid Rows

```{r missing-per-subject}
missing <- thew_1min %>%
  group_by(id, condition, dataset) %>%
  summarise(
    n_total      = n(),
    n_na_hr      = sum(is.na(avg_hr)),
    n_invalid    = sum(valid == FALSE, na.rm = TRUE),
    n_missing    = sum(is.na(avg_hr) | valid == FALSE),
    pct_na_hr    = n_na_hr / n_total * 100,
    pct_invalid  = n_invalid / n_total * 100,
    pct_missing  = n_missing / n_total * 100,
    .groups      = "drop"
  ) %>%
  arrange(desc(pct_missing))
```

```{r missing-summary}
missing %>%
  summarise(
    subjects_any_missing   = sum(pct_missing > 0),
    subjects_gt5pct        = sum(pct_missing > 5),
    subjects_gt20pct       = sum(pct_missing > 20),
    median_pct_missing     = median(pct_missing),
    max_pct_missing        = max(pct_missing)
  ) %>%
  kable(digits = 1, caption = "Missing data summary (NA or invalid)")
```

### Missingness by Condition

```{r missing-by-condition}
missing %>%
  group_by(condition) %>%
  summarise(
    n_subjects         = n(),
    median_pct_missing = median(pct_missing),
    mean_pct_invalid   = mean(pct_invalid),
    max_pct_missing    = max(pct_missing),
    .groups = "drop"
  ) %>%
  kable(digits = 1, caption = "Missingness by condition")
```

```{r missing-bar, fig.width=12, fig.height=5}
missing %>%
  mutate(id = fct_reorder(factor(id), pct_missing)) %>%
  ggplot(aes(x = id, y = pct_missing, fill = condition)) +
  geom_col() +
  labs(title = "Percent missing/invalid HR per subject",
       x = "Subject (sorted)", y = "% missing or invalid minutes") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

## 4 Descriptive Statistics

### Overall

```{r overall-stats}
thew_1min %>%
  filter(!is.na(avg_hr), valid == TRUE) %>%
  summarise(
    n_obs     = n(),
    mean_hr   = mean(avg_hr),
    sd_hr     = sd(avg_hr),
    median_hr = median(avg_hr),
    iqr_hr    = IQR(avg_hr),
    min_hr    = min(avg_hr),
    max_hr    = max(avg_hr)
  ) %>%
  kable(digits = 1, caption = "Overall HR descriptive statistics (1-min, valid only)")
```

### Per-Subject

```{r per-subject-stats}
subj_stats <- thew_1min %>%
  filter(!is.na(avg_hr), valid == TRUE) %>%
  group_by(id, condition, dataset) %>%
  summarise(
    mean_hr    = mean(avg_hr),
    sd_hr      = sd(avg_hr),
    min_hr     = min(avg_hr),
    max_hr     = max(avg_hr),
    n_valid    = n(),
    .groups    = "drop"
  )

subj_stats %>%
  head(20) %>%
  kable(digits = 1, caption = "Per-subject stats (first 20)")
```

### By Condition

```{r group-boxplot, fig.width=8, fig.height=5}
subj_stats %>%
  ggplot(aes(x = condition, y = mean_hr, fill = condition)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 21) +
  geom_jitter(width = 0.2, alpha = 0.3, size = 1) +
  labs(title = "Mean HR by condition",
       y = "Subject mean HR (bpm)") +
  theme(legend.position = "none")
```

### Sanity Checks

```{r sanity-checks}
flag_mean <- subj_stats %>% filter(mean_hr < 40 | mean_hr > 150)
flag_short <- subj_stats %>% filter(n_valid < 720)
flag_artifact <- thew_1min %>%
  filter(!is.na(avg_hr), valid == TRUE, (avg_hr < 20 | avg_hr > 250))

cat("Subjects with mean HR outside [40, 150] bpm:", nrow(flag_mean), "\n")
cat("Subjects with <12 hours valid data:", nrow(flag_short), "\n")
cat("Individual minutes with HR outside [20, 250] bpm:", nrow(flag_artifact), "\n")
```

```{r flag-tables}
if (nrow(flag_mean) > 0) {
  flag_mean %>%
    select(id, condition, mean_hr) %>%
    kable(digits = 1, caption = "Flagged subjects: mean HR outside [40, 150]")
}

if (nrow(flag_short) > 0) {
  flag_short %>%
    select(id, condition, dataset, n_valid, mean_hr) %>%
    head(20) %>%
    kable(digits = 1, caption = "Subjects with <12h valid data (first 20)")
}
```

## 5 HR Time Series Plots (1-min)

### Individual Traces (Sampled)

```{r sample-subjects-1min, fig.width=14, fig.height=12}
set.seed(42)

sampled_ids <- thew_1min %>%
  distinct(id, condition) %>%
  group_by(condition) %>%
  slice_sample(n = 6) %>%
  ungroup()

thew_1min %>%
  semi_join(sampled_ids, by = "id") %>%
  filter(valid == TRUE) %>%
  mutate(hours = t / 60) %>%
  ggplot(aes(x = hours, y = avg_hr, color = condition)) +
  geom_line(linewidth = 0.3) +
  facet_wrap(~ id, ncol = 3, scales = "free_y") +
  labs(title = "Sampled individual 24h HR traces (1-min)",
       x = "Time (hours)", y = "HR (bpm)") +
  theme(legend.position = "bottom")
```

### Spaghetti Plots by Condition

```{r spaghetti-1min, fig.width=14, fig.height=6}
group_means_1min <- thew_1min %>%
  filter(!is.na(avg_hr), valid == TRUE) %>%
  mutate(hours = t / 60) %>%
  group_by(condition, hours) %>%
  summarise(group_mean = mean(avg_hr, na.rm = TRUE), .groups = "drop")

thew_1min %>%
  filter(!is.na(avg_hr), valid == TRUE) %>%
  mutate(hours = t / 60) %>%
  ggplot(aes(x = hours, y = avg_hr, group = id)) +
  geom_line(alpha = 0.1, linewidth = 0.15) +
  geom_line(data = group_means_1min,
            aes(x = hours, y = group_mean, group = 1),
            color = "red", linewidth = 1) +
  facet_wrap(~ condition) +
  labs(title = "All subjects overlaid with group mean (1-min, red = mean)",
       x = "Time (hours)", y = "HR (bpm)")
```

## 6 HR Time Series Plots (5-min)

### Individual Traces (Sampled)

```{r sample-subjects-5min, fig.width=14, fig.height=12}
set.seed(42)

sampled_ids_5 <- thew_5min %>%
  distinct(id, condition) %>%
  group_by(condition) %>%
  slice_sample(n = 6) %>%
  ungroup()

thew_5min %>%
  semi_join(sampled_ids_5, by = "id") %>%
  filter(valid == TRUE) %>%
  mutate(hours = t / 60) %>%
  ggplot(aes(x = hours, y = avg_hr, color = condition)) +
  geom_line(linewidth = 0.4) +
  facet_wrap(~ id, ncol = 3, scales = "free_y") +
  labs(title = "Sampled individual 24h HR traces (5-min)",
       x = "Time (hours)", y = "HR (bpm)") +
  theme(legend.position = "bottom")
```

### Spaghetti Plots by Condition

```{r spaghetti-5min, fig.width=14, fig.height=6}
group_means_5min <- thew_5min %>%
  filter(!is.na(avg_hr), valid == TRUE) %>%
  mutate(hours = t / 60) %>%
  group_by(condition, hours) %>%
  summarise(group_mean = mean(avg_hr, na.rm = TRUE), .groups = "drop")

thew_5min %>%
  filter(!is.na(avg_hr), valid == TRUE) %>%
  mutate(hours = t / 60) %>%
  ggplot(aes(x = hours, y = avg_hr, group = id)) +
  geom_line(alpha = 0.1, linewidth = 0.15) +
  geom_line(data = group_means_5min,
            aes(x = hours, y = group_mean, group = 1),
            color = "red", linewidth = 1) +
  facet_wrap(~ condition) +
  labs(title = "All subjects overlaid with group mean (5-min, red = mean)",
       x = "Time (hours)", y = "HR (bpm)")
```

## 7 Distribution Plots

### Density of All HR Values (1-min)

```{r density-1min, fig.width=10, fig.height=5}
thew_1min %>%
  filter(!is.na(avg_hr), valid == TRUE) %>%
  ggplot(aes(x = avg_hr, fill = condition)) +
  geom_density(alpha = 0.4) +
  labs(title = "Density of 1-min HR values by condition",
       x = "HR (bpm)", y = "Density") +
  scale_x_continuous(limits = c(30, 180))
```

### Per-Subject Boxplots (Sampled)

```{r subject-boxplots, fig.width=14, fig.height=6}
set.seed(42)
boxplot_ids <- thew_1min %>%
  distinct(id, condition) %>%
  group_by(condition) %>%
  slice_sample(n = 8) %>%
  ungroup()

thew_1min %>%
  semi_join(boxplot_ids, by = "id") %>%
  filter(!is.na(avg_hr), valid == TRUE) %>%
  mutate(id = factor(id)) %>%
  ggplot(aes(x = id, y = avg_hr, fill = condition)) +
  geom_boxplot(outlier.size = 0.5) +
  labs(title = "Per-subject HR distributions (sampled, 1-min)",
       x = "Subject", y = "HR (bpm)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 8 Full Subject Gallery

Every subject's 1-min HR trace, faceted by ID and colored by condition.

```{r all-subjects-gallery, fig.width=24, fig.height=80}
thew_1min %>%
  filter(!is.na(avg_hr), valid == TRUE) %>%
  mutate(hours = t / 60) %>%
  ggplot(aes(x = hours, y = avg_hr, color = condition)) +
  geom_line(linewidth = 0.15) +
  facet_wrap(~ id, ncol = 12) +
  labs(title = "All subjects — 24h HR traces (1-min)",
       x = "Hours", y = "HR (bpm)") +
  theme_minimal(base_size = 6) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(size = 4),
    axis.text = element_text(size = 3),
    panel.spacing = unit(0.15, "lines")
  )
```
