---
title: "EDA: PhysioNet CHAOS HR Time Series"
subtitle: "Exploratory data analysis of 1-min and 5-min downsampled heart rate (AF, CHF, Normal)"
date: today
format:
  gfm:
    toc: true
    toc-depth: 3
  html:
    self-contained: true
    toc: true
    toc-depth: 3
    code-fold: true
execute:
  warning: false
  message: false
  echo: false
---

## 1 Setup & Data Loading

```{r setup}
library(tidyverse)
library(here)
library(knitr)
library(scales)

theme_set(theme_minimal(base_size = 12))
```

```{r load-data}
pn_1min <- read_csv(here("data_processed/physionet_15/physionet15_1min.csv"),
                    show_col_types = FALSE)
pn_5min <- read_csv(here("data_processed/physionet_15/physionet15_5min.csv"),
                    show_col_types = FALSE)
```

```{r data-overview}
cat("--- 1-minute data ---\n")
cat("Dimensions:", nrow(pn_1min), "rows x", ncol(pn_1min), "cols\n")
cat("Subjects:", n_distinct(pn_1min$id), "\n")
cat("Conditions:", paste(unique(pn_1min$condition), collapse = ", "), "\n\n")

cat("--- 5-minute data ---\n")
cat("Dimensions:", nrow(pn_5min), "rows x", ncol(pn_5min), "cols\n")
cat("Subjects:", n_distinct(pn_5min$id), "\n")
```

```{r glimpse}
glimpse(pn_1min)
```

## 2 Subject-Level Summary

### Counts by Condition

```{r condition-counts}
pn_1min %>%
  distinct(id, condition, source_record) %>%
  count(condition, name = "n_subjects") %>%
  kable(caption = "Subjects by condition")
```

### Subject Inventory

```{r subject-inventory}
pn_1min %>%
  distinct(id, condition, source_record) %>%
  arrange(condition, id) %>%
  kable(caption = "All subjects with source records")
```

### Recording Duration

```{r recording-duration}
dur <- pn_1min %>%
  group_by(id, condition) %>%
  summarise(
    n_minutes   = n(),
    n_valid_min = sum(!is.na(mean_hr)),
    duration_hr = n_minutes / 60,
    .groups = "drop"
  )

dur %>%
  group_by(condition) %>%
  summarise(
    n_subjects       = n(),
    median_minutes   = median(n_minutes),
    min_minutes      = min(n_minutes),
    max_minutes      = max(n_minutes),
    median_valid_min = median(n_valid_min),
    .groups = "drop"
  ) %>%
  kable(caption = "Recording duration by condition")
```

```{r duration-per-subject}
dur %>%
  arrange(condition, id) %>%
  select(id, condition, n_minutes, n_valid_min, duration_hr) %>%
  kable(digits = 1, caption = "Per-subject recording duration")
```

## 3 Missing Data Audit

```{r missing-per-subject}
missing <- pn_1min %>%
  group_by(id, condition) %>%
  summarise(
    n_total = n(),
    n_na    = sum(is.na(mean_hr)),
    pct_na  = n_na / n_total * 100,
    .groups = "drop"
  ) %>%
  arrange(desc(pct_na))
```

```{r missing-summary}
missing %>%
  summarise(
    subjects_any_missing = sum(n_na > 0),
    subjects_gt5pct      = sum(pct_na > 5),
    subjects_gt20pct     = sum(pct_na > 20),
    median_pct_missing   = median(pct_na),
    max_pct_missing      = max(pct_na)
  ) %>%
  kable(digits = 1, caption = "Missing data summary (mean_hr NAs)")
```

```{r missing-bar, fig.width=8, fig.height=4}
missing %>%
  mutate(id = fct_reorder(factor(id), pct_na)) %>%
  ggplot(aes(x = id, y = pct_na, fill = condition)) +
  geom_col() +
  labs(title = "Percent missing HR per subject",
       x = "Subject", y = "% missing minutes") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 4 Descriptive Statistics

### Overall

```{r overall-stats}
pn_1min %>%
  filter(!is.na(mean_hr)) %>%
  summarise(
    n_obs     = n(),
    mean_hr   = mean(mean_hr),
    sd_hr     = sd(mean_hr),
    median_hr = median(mean_hr),
    iqr_hr    = IQR(mean_hr),
    min_hr    = min(mean_hr),
    max_hr    = max(mean_hr)
  ) %>%
  kable(digits = 1, caption = "Overall HR descriptive statistics (1-min)")
```

### Per-Subject

```{r per-subject-stats}
subj_stats <- pn_1min %>%
  filter(!is.na(mean_hr)) %>%
  group_by(id, condition) %>%
  summarise(
    mean_hr  = mean(mean_hr),
    sd_hr    = sd(mean_hr),
    min_hr   = min(mean_hr),
    max_hr   = max(mean_hr),
    n_valid  = n(),
    .groups  = "drop"
  )

subj_stats %>%
  arrange(condition, id) %>%
  kable(digits = 1, caption = "Per-subject HR statistics")
```

### By Condition

```{r group-boxplot, fig.width=7, fig.height=5}
subj_stats %>%
  ggplot(aes(x = condition, y = mean_hr, fill = condition)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 21) +
  geom_jitter(width = 0.15, alpha = 0.6, size = 2.5) +
  labs(title = "Mean HR by condition",
       y = "Subject mean HR (bpm)") +
  theme(legend.position = "none")
```

### Sanity Checks

```{r sanity-checks}
flag_mean <- subj_stats %>% filter(mean_hr < 40 | mean_hr > 150)
flag_short <- subj_stats %>% filter(n_valid < 720)
flag_artifact <- pn_1min %>% filter(!is.na(mean_hr), mean_hr < 20 | mean_hr > 250)

cat("Subjects with mean HR outside [40, 150] bpm:", nrow(flag_mean), "\n")
cat("Subjects with <12 hours valid data:", nrow(flag_short), "\n")
cat("Individual minutes with HR outside [20, 250] bpm:", nrow(flag_artifact), "\n")
```

```{r flag-tables}
if (nrow(flag_mean) > 0) {
  flag_mean %>%
    select(id, condition, mean_hr) %>%
    kable(digits = 1, caption = "Flagged subjects: mean HR outside [40, 150]")
}

if (nrow(flag_short) > 0) {
  flag_short %>%
    select(id, condition, n_valid, mean_hr) %>%
    kable(digits = 1, caption = "Subjects with <12h valid data")
}
```

## 5 HR Time Series Plots (1-min)

### All Individual Traces

With only 15 subjects we can show every trace individually.

```{r all-subjects-1min, fig.width=14, fig.height=10}
pn_1min %>%
  filter(!is.na(mean_hr)) %>%
  mutate(hours = minute_index / 60) %>%
  ggplot(aes(x = hours, y = mean_hr, color = condition)) +
  geom_line(linewidth = 0.3) +
  facet_wrap(~ id, ncol = 5, scales = "free_y") +
  labs(title = "All subjects — 24h HR traces (1-min)",
       x = "Time (hours)", y = "HR (bpm)") +
  theme(legend.position = "bottom")
```

### Spaghetti Plots by Condition

```{r spaghetti-1min, fig.width=14, fig.height=5}
group_means_1min <- pn_1min %>%
  filter(!is.na(mean_hr)) %>%
  mutate(hours = minute_index / 60) %>%
  group_by(condition, hours) %>%
  summarise(group_mean = mean(mean_hr, na.rm = TRUE), .groups = "drop")

pn_1min %>%
  filter(!is.na(mean_hr)) %>%
  mutate(hours = minute_index / 60) %>%
  ggplot(aes(x = hours, y = mean_hr, group = id)) +
  geom_line(alpha = 0.3, linewidth = 0.3) +
  geom_line(data = group_means_1min,
            aes(x = hours, y = group_mean, group = 1),
            color = "red", linewidth = 1) +
  facet_wrap(~ condition) +
  labs(title = "All subjects overlaid with group mean (1-min, red = mean)",
       x = "Time (hours)", y = "HR (bpm)")
```

## 6 HR Time Series Plots (5-min)

### All Individual Traces

```{r all-subjects-5min, fig.width=14, fig.height=10}
pn_5min %>%
  filter(!is.na(mean_hr)) %>%
  mutate(hours = minute_index / 60) %>%
  ggplot(aes(x = hours, y = mean_hr, color = condition)) +
  geom_line(linewidth = 0.4) +
  facet_wrap(~ id, ncol = 5, scales = "free_y") +
  labs(title = "All subjects — 24h HR traces (5-min)",
       x = "Time (hours)", y = "HR (bpm)") +
  theme(legend.position = "bottom")
```

### Spaghetti Plots by Condition

```{r spaghetti-5min, fig.width=14, fig.height=5}
group_means_5min <- pn_5min %>%
  filter(!is.na(mean_hr)) %>%
  mutate(hours = minute_index / 60) %>%
  group_by(condition, hours) %>%
  summarise(group_mean = mean(mean_hr, na.rm = TRUE), .groups = "drop")

pn_5min %>%
  filter(!is.na(mean_hr)) %>%
  mutate(hours = minute_index / 60) %>%
  ggplot(aes(x = hours, y = mean_hr, group = id)) +
  geom_line(alpha = 0.3, linewidth = 0.3) +
  geom_line(data = group_means_5min,
            aes(x = hours, y = group_mean, group = 1),
            color = "red", linewidth = 1) +
  facet_wrap(~ condition) +
  labs(title = "All subjects overlaid with group mean (5-min, red = mean)",
       x = "Time (hours)", y = "HR (bpm)")
```

## 7 Distribution Plots

### Density of All HR Values (1-min)

```{r density-1min, fig.width=10, fig.height=5}
pn_1min %>%
  filter(!is.na(mean_hr)) %>%
  ggplot(aes(x = mean_hr, fill = condition)) +
  geom_density(alpha = 0.4) +
  labs(title = "Density of 1-min HR values by condition",
       x = "HR (bpm)", y = "Density") +
  scale_x_continuous(limits = c(30, 200))
```

### Per-Subject Boxplots

```{r subject-boxplots, fig.width=12, fig.height=5}
pn_1min %>%
  filter(!is.na(mean_hr)) %>%
  mutate(id = fct_reorder(factor(id), mean_hr, .fun = median)) %>%
  ggplot(aes(x = id, y = mean_hr, fill = condition)) +
  geom_boxplot(outlier.size = 0.5) +
  labs(title = "Per-subject HR distributions (1-min)",
       x = "Subject (sorted by median HR)", y = "HR (bpm)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
