---
title: "EDA: SHAREE HR Time Series"
subtitle: "Exploratory data analysis of 1-min and 5-min downsampled heart rate"
date: today
format:
  gfm:
    toc: true
    toc-depth: 3
  html:
    self-contained: true
    toc: true
    toc-depth: 3
    code-fold: true
execute:
  warning: false
  message: false
  echo: false
---

## 1 Setup & Data Loading

```{r setup}
library(tidyverse)
library(here)
library(knitr)
library(scales)

theme_set(theme_minimal(base_size = 12))
```

```{r load-data}
sharee_1min <- read_csv(here("data_processed/sharee_20251118/sharee_1min.csv"),
                        show_col_types = FALSE)
sharee_5min <- read_csv(here("data_processed/sharee_20251118/sharee_5min.csv"),
                        show_col_types = FALSE)
```

```{r data-overview}
cat("--- 1-minute data ---\n")
cat("Dimensions:", nrow(sharee_1min), "rows x", ncol(sharee_1min), "cols\n")
cat("Subjects:", n_distinct(sharee_1min$id), "\n\n")

cat("--- 5-minute data ---\n")
cat("Dimensions:", nrow(sharee_5min), "rows x", ncol(sharee_5min), "cols\n")
cat("Subjects:", n_distinct(sharee_5min$id), "\n")
```

```{r glimpse}
glimpse(sharee_1min)
```

## 2 Subject-Level Summary

### Demographics

```{r demographics}
demo <- sharee_1min %>%
  distinct(id, .keep_all = TRUE) %>%
  select(id, Gender, Age, BMI, Smoker, SBP, DBP, Hypertension)

demo %>%
  summarise(
    n = n(),
    age_mean = mean(Age, na.rm = TRUE),
    age_sd   = sd(Age, na.rm = TRUE),
    bmi_mean = mean(BMI, na.rm = TRUE),
    bmi_sd   = sd(BMI, na.rm = TRUE),
    pct_male = mean(Gender == "M", na.rm = TRUE) * 100,
    pct_smoker = mean(Smoker == "yes", na.rm = TRUE) * 100
  ) %>%
  kable(digits = 1, caption = "Overall demographics")
```

### Counts by Hypertension Category

```{r htn-counts}
demo %>%
  count(Hypertension, name = "n_subjects") %>%
  kable(caption = "Subjects by hypertension category")
```

### Recording Duration

```{r recording-duration}
dur <- sharee_1min %>%
  group_by(id, Hypertension) %>%
  summarise(
    n_minutes   = n(),
    n_valid_min = sum(!is.na(mean_hr)),
    duration_hr = n_minutes / 60,
    .groups = "drop"
  )

dur %>%
  summarise(
    n_subjects       = n(),
    median_minutes   = median(n_minutes),
    min_minutes      = min(n_minutes),
    max_minutes      = max(n_minutes),
    median_valid_min = median(n_valid_min)
  ) %>%
  kable(caption = "Recording duration summary")
```

## 3 Missing Data Audit

```{r missing-per-subject}
missing <- sharee_1min %>%
  group_by(id, Hypertension) %>%
  summarise(
    n_total  = n(),
    n_na     = sum(is.na(mean_hr)),
    pct_na   = n_na / n_total * 100,
    .groups  = "drop"
  ) %>%
  arrange(desc(pct_na))
```

```{r missing-summary}
missing %>%
  summarise(
    subjects_any_missing   = sum(n_na > 0),
    subjects_gt5pct        = sum(pct_na > 5),
    subjects_gt20pct       = sum(pct_na > 20),
    median_pct_missing     = median(pct_na),
    max_pct_missing        = max(pct_na)
  ) %>%
  kable(digits = 1, caption = "Missing data summary (mean_hr NAs)")
```

```{r missing-bar, fig.width=10, fig.height=5}
missing %>%
  mutate(id = fct_reorder(factor(id), pct_na)) %>%
  ggplot(aes(x = id, y = pct_na, fill = Hypertension)) +
  geom_col() +
  labs(title = "Percent missing HR per subject",
       x = "Subject (sorted)", y = "% missing minutes") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

## 4 Descriptive Statistics

### Overall

```{r overall-stats}
sharee_1min %>%
  filter(!is.na(mean_hr)) %>%
  summarise(
    n_obs    = n(),
    mean_hr  = mean(mean_hr),
    sd_hr    = sd(mean_hr),
    median_hr = median(mean_hr),
    iqr_hr   = IQR(mean_hr),
    min_hr   = min(mean_hr),
    max_hr   = max(mean_hr)
  ) %>%
  kable(digits = 1, caption = "Overall HR descriptive statistics (1-min)")
```

### Per-Subject

```{r per-subject-stats}
subj_stats <- sharee_1min %>%
  filter(!is.na(mean_hr)) %>%
  group_by(id, Hypertension) %>%
  summarise(
    mean_hr    = mean(mean_hr),
    sd_hr      = sd(mean_hr),
    min_hr     = min(mean_hr),
    max_hr     = max(mean_hr),
    n_valid    = n(),
    .groups    = "drop"
  )

subj_stats %>%
  head(20) %>%
  kable(digits = 1, caption = "Per-subject stats (first 20)")
```

### By Hypertension Category

```{r group-boxplot, fig.width=8, fig.height=5}
subj_stats %>%
  ggplot(aes(x = Hypertension, y = mean_hr, fill = Hypertension)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 21) +
  geom_jitter(width = 0.15, alpha = 0.4, size = 1.5) +
  labs(title = "Mean HR by hypertension category",
       y = "Subject mean HR (bpm)") +
  theme(legend.position = "none")
```

### Sanity Checks

```{r sanity-checks}
flag_mean <- subj_stats %>% filter(mean_hr < 40 | mean_hr > 150)
flag_short <- subj_stats %>% filter(n_valid < 720)
flag_artifact <- sharee_1min %>% filter(!is.na(mean_hr), mean_hr < 20 | mean_hr > 250)

cat("Subjects with mean HR outside [40, 150] bpm:", nrow(flag_mean), "\n")
cat("Subjects with <12 hours valid data:", nrow(flag_short), "\n")
cat("Individual minutes with HR outside [20, 250] bpm:", nrow(flag_artifact), "\n")
```

```{r flag-tables}
if (nrow(flag_mean) > 0) {
  cat("Flagged subjects (mean HR):\n")
  flag_mean %>% select(id, Hypertension, mean_hr) %>% kable(digits = 1)
}

if (nrow(flag_short) > 0) {
  cat("Flagged subjects (<12h valid):\n")
  flag_short %>% select(id, Hypertension, n_valid, mean_hr) %>%
    head(20) %>%
    kable(digits = 1, caption = "Subjects with <12h valid data (first 20)")
}
```

## 5 HR Time Series Plots (1-min)

### Individual Traces (Sampled)

```{r sample-subjects-1min, fig.width=12, fig.height=10}
set.seed(42)

sampled_ids <- sharee_1min %>%
  distinct(id, Hypertension) %>%
  group_by(Hypertension) %>%
  slice_sample(n = 4) %>%
  ungroup()

sharee_1min %>%
  semi_join(sampled_ids, by = "id") %>%
  mutate(hours = minute_index / 60) %>%
  ggplot(aes(x = hours, y = mean_hr, color = Hypertension)) +
  geom_line(linewidth = 0.3) +
  facet_wrap(~ id, ncol = 4, scales = "free_y") +
  labs(title = "Sampled individual 24h HR traces (1-min)",
       x = "Time (hours)", y = "HR (bpm)") +
  theme(legend.position = "bottom")
```

### Spaghetti Plots by Group

```{r spaghetti-1min, fig.width=12, fig.height=8}
group_means_1min <- sharee_1min %>%
  filter(!is.na(mean_hr)) %>%
  mutate(hours = minute_index / 60) %>%
  group_by(Hypertension, hours) %>%
  summarise(group_mean = mean(mean_hr, na.rm = TRUE), .groups = "drop")

sharee_1min %>%
  filter(!is.na(mean_hr)) %>%
  mutate(hours = minute_index / 60) %>%
  ggplot(aes(x = hours, y = mean_hr, group = id)) +
  geom_line(alpha = 0.15, linewidth = 0.2) +
  geom_line(data = group_means_1min,
            aes(x = hours, y = group_mean, group = 1),
            color = "red", linewidth = 1) +
  facet_wrap(~ Hypertension) +
  labs(title = "All subjects overlaid with group mean (1-min, red = mean)",
       x = "Time (hours)", y = "HR (bpm)")
```

## 6 HR Time Series Plots (5-min)

### Individual Traces (Sampled)

```{r sample-subjects-5min, fig.width=12, fig.height=10}
set.seed(42)

sampled_ids_5 <- sharee_5min %>%
  distinct(id, Hypertension) %>%
  group_by(Hypertension) %>%
  slice_sample(n = 4) %>%
  ungroup()

sharee_5min %>%
  semi_join(sampled_ids_5, by = "id") %>%
  mutate(hours = minute_index / 60) %>%
  ggplot(aes(x = hours, y = mean_hr, color = Hypertension)) +
  geom_line(linewidth = 0.4) +
  facet_wrap(~ id, ncol = 4, scales = "free_y") +
  labs(title = "Sampled individual 24h HR traces (5-min)",
       x = "Time (hours)", y = "HR (bpm)") +
  theme(legend.position = "bottom")
```

### Spaghetti Plots by Group

```{r spaghetti-5min, fig.width=12, fig.height=8}
group_means_5min <- sharee_5min %>%
  filter(!is.na(mean_hr)) %>%
  mutate(hours = minute_index / 60) %>%
  group_by(Hypertension, hours) %>%
  summarise(group_mean = mean(mean_hr, na.rm = TRUE), .groups = "drop")

sharee_5min %>%
  filter(!is.na(mean_hr)) %>%
  mutate(hours = minute_index / 60) %>%
  ggplot(aes(x = hours, y = mean_hr, group = id)) +
  geom_line(alpha = 0.15, linewidth = 0.2) +
  geom_line(data = group_means_5min,
            aes(x = hours, y = group_mean, group = 1),
            color = "red", linewidth = 1) +
  facet_wrap(~ Hypertension) +
  labs(title = "All subjects overlaid with group mean (5-min, red = mean)",
       x = "Time (hours)", y = "HR (bpm)")
```

## 7 Distribution Plots

### Density of All HR Values (1-min)

```{r density-1min, fig.width=10, fig.height=5}
sharee_1min %>%
  filter(!is.na(mean_hr)) %>%
  ggplot(aes(x = mean_hr, fill = Hypertension)) +
  geom_density(alpha = 0.4) +
  labs(title = "Density of 1-min HR values by hypertension category",
       x = "HR (bpm)", y = "Density") +
  scale_x_continuous(limits = c(30, 180))
```

### Per-Subject Boxplots (Sampled)

```{r subject-boxplots, fig.width=12, fig.height=6}
set.seed(42)
boxplot_ids <- sharee_1min %>%
  distinct(id, Hypertension) %>%
  group_by(Hypertension) %>%
  slice_sample(n = 6) %>%
  ungroup()

sharee_1min %>%
  semi_join(boxplot_ids, by = "id") %>%
  filter(!is.na(mean_hr)) %>%
  mutate(id = factor(id)) %>%
  ggplot(aes(x = id, y = mean_hr, fill = Hypertension)) +
  geom_boxplot(outlier.size = 0.5) +
  labs(title = "Per-subject HR distributions (sampled, 1-min)",
       x = "Subject", y = "HR (bpm)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 8 Full Subject Gallery

Every subject's 1-min HR trace, faceted by ID and colored by hypertension category.

```{r all-subjects-gallery, fig.width=20, fig.height=40}
sharee_1min %>%
  filter(!is.na(mean_hr)) %>%
  mutate(hours = minute_index / 60) %>%
  ggplot(aes(x = hours, y = mean_hr, color = Hypertension)) +
  geom_line(linewidth = 0.2) +
  facet_wrap(~ id, ncol = 10) +
  labs(title = "All subjects — 24h HR traces (1-min)",
       x = "Hours", y = "HR (bpm)") +
  theme_minimal(base_size = 6) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(size = 5),
    axis.text = element_text(size = 4),
    panel.spacing = unit(0.2, "lines")
  )
```
