---
title: "THEW Holter Metrics: Group Comparisons by Health Status"
subtitle: "Catch22 complexity, entropy, and HRV metrics across CAD, CAD+Diabetes, ESRD, and Healthy cohorts"
date: today
format:
  gfm:
    toc: true
    toc-depth: 3
  html:
    self-contained: true
    toc: true
    toc-depth: 3
    code-fold: true
execute:
  warning: false
  message: false
  echo: false
---

## 1 Setup & Data Loading

```{r setup}
library(tidyverse)
library(here)
library(knitr)
library(broom)
library(scales)

theme_set(theme_minimal(base_size = 12))
set.seed(42)
```

```{r load-data}
metrics_1min <- read_csv(here("results/thew/thew_metrics_1min.csv"),
                         show_col_types = FALSE)
metrics_5min <- read_csv(here("results/thew/thew_metrics_5min.csv"),
                         show_col_types = FALSE)
```

```{r define-columns}
# --- Metric column definitions ---
# 24 catch22 + 7 descriptive stats + 3 entropy = 34 avg_hr columns
avg_hr_cols <- c(

  # catch22 (24)
  "avg_hr_DN_HistogramMode_5", "avg_hr_DN_HistogramMode_10",
  "avg_hr_CO_f1ecac", "avg_hr_CO_FirstMin_ac",
  "avg_hr_CO_HistogramAMI_even_2_5", "avg_hr_CO_trev_1_num",
  "avg_hr_MD_hrv_classic_pnn40",
  "avg_hr_SB_BinaryStats_mean_longstretch1",
  "avg_hr_SB_TransitionMatrix_3ac_sumdiagcov",
  "avg_hr_PD_PeriodicityWang_th0_01",
  "avg_hr_CO_Embed2_Dist_tau_d_expfit_meandiff",
  "avg_hr_IN_AutoMutualInfoStats_40_gaussian_fmmi",
  "avg_hr_FC_LocalSimple_mean1_tauresrat",
  "avg_hr_DN_OutlierInclude_p_001_mdrmd",
  "avg_hr_DN_OutlierInclude_n_001_mdrmd",
  "avg_hr_SP_Summaries_welch_rect_area_5_1",
  "avg_hr_SB_BinaryStats_diff_longstretch0",
  "avg_hr_SB_MotifThree_quantile_hh",
  "avg_hr_SC_FluctAnal_2_rsrangefit_50_1_logi_prop_r1",
  "avg_hr_SC_FluctAnal_2_dfa_50_1_2_logi_prop_r1",
  "avg_hr_SP_Summaries_welch_rect_centroid",
  "avg_hr_FC_LocalSimple_mean3_stderr",
  "avg_hr_DN_Mean", "avg_hr_DN_Spread_Std",
  # descriptive stats (7)
  "avg_hr_mean", "avg_hr_sd", "avg_hr_max", "avg_hr_min",
  "avg_hr_cv", "avg_hr_iqr", "avg_hr_median",
  # entropy (3)
  "avg_hr_apen", "avg_hr_sampen", "avg_hr_permen"
)

# HRV columns (4) — mean_beat_count excluded (data-quantity, not HR dynamics)
hrv_cols <- c("mean_sdnn_ms", "mean_rmssd_ms", "mean_median_hr", "mean_sd_hr")

all_metric_cols <- c(avg_hr_cols, hrv_cols)

# Guardrails
stopifnot(length(avg_hr_cols) == 34)
stopifnot(length(hrv_cols) == 4)
stopifnot(length(all_metric_cols) == 38)
```

```{r prepare-data}
# Factor condition4 with explicit level order
condition4_levels <- c("Healthy", "CAD", "CAD_Diabetes", "ESRD")

prepare_data <- function(df) {
  stopifnot(all(c("condition4", all_metric_cols) %in% names(df)))
  df %>%
    mutate(condition4 = factor(condition4, levels = condition4_levels))
}

metrics_1min <- prepare_data(metrics_1min)
metrics_5min <- prepare_data(metrics_5min)

# Color palette for the 4 groups
condition4_colors <- c(
  "Healthy"      = "#4DAF4A",
  "CAD"          = "#377EB8",
  "CAD_Diabetes" = "#FF7F00",
  "ESRD"         = "#E41A1C"
)
```


## 2 Data Overview & Demographics

```{r group-sizes}
bind_rows(
  metrics_5min %>% count(condition4, name = "n") %>% mutate(resolution = "5-min"),
  metrics_1min %>% count(condition4, name = "n") %>% mutate(resolution = "1-min")
) %>%
  pivot_wider(names_from = resolution, values_from = n) %>%
  kable(caption = "Subjects by condition4 group")
```

```{r demographics}
demo_cols <- c("AGE", "BMI", "GENDER")
demo_available <- intersect(demo_cols, names(metrics_5min))

if (length(demo_available) > 0) {
  # Numeric demographics
  num_demo <- intersect(c("AGE", "BMI"), demo_available)
  if (length(num_demo) > 0) {
    metrics_5min %>%
      group_by(condition4) %>%
      summarize(
        across(all_of(num_demo),
               list(mean = ~mean(.x, na.rm = TRUE),
                    sd   = ~sd(.x, na.rm = TRUE),
                    n_NA = ~sum(is.na(.x))),
               .names = "{.col}_{.fn}"),
        .groups = "drop"
      ) %>%
      kable(digits = 1, caption = "Demographics by condition4 (mean +/- SD)")
  }

  # Gender breakdown
  if ("GENDER" %in% demo_available) {
    metrics_5min %>%
      count(condition4, GENDER) %>%
      pivot_wider(names_from = GENDER, values_from = n, values_fill = 0) %>%
      kable(caption = "Gender distribution by condition4")
  }
}
```

```{r na-audit}
na_counts <- metrics_5min %>%
  summarize(across(all_of(all_metric_cols), ~sum(is.na(.x)))) %>%
  pivot_longer(everything(), names_to = "metric", values_to = "n_NA_5min")

na_counts_1 <- metrics_1min %>%
  summarize(across(all_of(all_metric_cols), ~sum(is.na(.x)))) %>%
  pivot_longer(everything(), names_to = "metric", values_to = "n_NA_1min")

na_summary <- left_join(na_counts, na_counts_1, by = "metric") %>%
  filter(n_NA_5min > 0 | n_NA_1min > 0)

if (nrow(na_summary) > 0) {
  na_summary %>% kable(caption = "Metrics with missing values")
} else {
  cat("No missing values in any metric column.\n")
}
```


## 3 Boxplots by Condition

```{r boxplot-helpers}
# Helper to pivot metrics long and strip avg_hr_ prefix for labels
pivot_metrics_long <- function(df, metric_cols) {
  df %>%
    select(id, condition4, all_of(metric_cols)) %>%
    pivot_longer(
      cols = all_of(metric_cols),
      names_to = "metric",
      values_to = "value"
    ) %>%
    mutate(metric_label = str_remove(metric, "^avg_hr_"))
}

# Boxplot function
plot_boxplots <- function(df_long, ncol = 4, title = "") {
  ggplot(df_long, aes(x = condition4, y = value, fill = condition4)) +
    geom_boxplot(alpha = 0.7, outlier.shape = 21, outlier.size = 1) +
    facet_wrap(~ metric_label, scales = "free_y", ncol = ncol) +
    scale_fill_manual(values = condition4_colors) +
    labs(title = title, x = NULL, y = NULL, fill = "Condition") +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      strip.text = element_text(size = 7),
      legend.position = "bottom"
    )
}
```

### avg_hr Metrics (catch22 + descriptive + entropy)

::: {.panel-tabset}

#### 5-min

```{r boxplots-avghr-5min, fig.width=14, fig.height=22}
pivot_metrics_long(metrics_5min, avg_hr_cols) %>%
  plot_boxplots(ncol = 4, title = "avg_hr Metrics by Condition (5-min)")
```

#### 1-min

```{r boxplots-avghr-1min, fig.width=14, fig.height=22}
pivot_metrics_long(metrics_1min, avg_hr_cols) %>%
  plot_boxplots(ncol = 4, title = "avg_hr Metrics by Condition (1-min)")
```

:::

### HRV Metrics

::: {.panel-tabset}

#### 5-min

```{r boxplots-hrv-5min, fig.width=10, fig.height=5}
pivot_metrics_long(metrics_5min, hrv_cols) %>%
  plot_boxplots(ncol = 2, title = "HRV Metrics by Condition (5-min)")
```

#### 1-min

```{r boxplots-hrv-1min, fig.width=10, fig.height=5}
pivot_metrics_long(metrics_1min, hrv_cols) %>%
  plot_boxplots(ncol = 2, title = "HRV Metrics by Condition (1-min)")
```

:::


## 4 Omnibus Significance Tests (ANOVA & Kruskal-Wallis)

```{r omnibus-helper}
run_omnibus_tests <- function(df, metric_cols) {
  df_long <- pivot_metrics_long(df, metric_cols)

  df_long %>%
    group_by(metric, metric_label) %>%
    group_modify(~ {
      dat <- .x %>% filter(!is.na(value))
      if (n_distinct(dat$condition4) < 2) {
        return(tibble(
          test = c("ANOVA", "Kruskal-Wallis"),
          statistic = NA_real_, p.value = NA_real_
        ))
      }

      # ANOVA
      aov_res <- aov(value ~ condition4, data = dat)
      aov_tidy <- tidy(aov_res) %>%
        filter(term == "condition4") %>%
        select(statistic, p.value) %>%
        mutate(test = "ANOVA")

      # Kruskal-Wallis
      kw_res <- kruskal.test(value ~ condition4, data = dat)
      kw_tidy <- tidy(kw_res) %>%
        select(statistic, p.value) %>%
        mutate(test = "Kruskal-Wallis")

      bind_rows(aov_tidy, kw_tidy)
    }) %>%
    ungroup() %>%
    group_by(test) %>%
    mutate(p_adj = p.adjust(p.value, method = "BH")) %>%
    ungroup() %>%
    mutate(
      sig = case_when(
        p_adj < 0.001 ~ "***",
        p_adj < 0.01  ~ "**",
        p_adj < 0.05  ~ "*",
        TRUE          ~ ""
      )
    )
}

format_omnibus_table <- function(omnibus_df) {
  omnibus_df %>%
    select(metric_label, test, statistic, p.value, p_adj, sig) %>%
    mutate(
      statistic = round(statistic, 2),
      p.value = signif(p.value, 3),
      p_adj = signif(p_adj, 3)
    ) %>%
    arrange(p.value) %>%
    kable(col.names = c("Metric", "Test", "Statistic", "p-value", "p (BH-adj)", "Sig"))
}
```

::: {.panel-tabset}

### 5-min

```{r omnibus-5min}
omnibus_5min <- run_omnibus_tests(metrics_5min, all_metric_cols)
format_omnibus_table(omnibus_5min)
```

### 1-min

```{r omnibus-1min}
omnibus_1min <- run_omnibus_tests(metrics_1min, all_metric_cols)
format_omnibus_table(omnibus_1min)
```

:::


## 5 Pairwise Significance Tests

```{r pairwise-helper}
run_pairwise_tests <- function(df, metric_cols) {
  df_long <- pivot_metrics_long(df, metric_cols)
  pairs <- combn(condition4_levels, 2, simplify = FALSE)

  results <- map_dfr(pairs, function(pair) {
    df_pair <- df_long %>% filter(condition4 %in% pair, !is.na(value))

    df_pair %>%
      group_by(metric, metric_label) %>%
      group_modify(~ {
        dat <- .x
        if (n_distinct(dat$condition4) < 2 || nrow(dat) < 4) {
          return(tibble(
            group1 = pair[1], group2 = pair[2],
            t_stat = NA_real_, t_p = NA_real_,
            w_stat = NA_real_, w_p = NA_real_
          ))
        }

        # Welch's t-test
        tt <- tryCatch(
          t.test(value ~ condition4, data = dat),
          error = function(e) NULL
        )
        # Wilcoxon rank-sum
        wt <- tryCatch(
          wilcox.test(value ~ condition4, data = dat),
          error = function(e) NULL
        )

        tibble(
          group1 = pair[1], group2 = pair[2],
          t_stat = if (!is.null(tt)) tt$statistic else NA_real_,
          t_p    = if (!is.null(tt)) tt$p.value   else NA_real_,
          w_stat = if (!is.null(wt)) wt$statistic else NA_real_,
          w_p    = if (!is.null(wt)) wt$p.value   else NA_real_
        )
      }) %>%
      ungroup()
  })

  # BH adjustment across all pairwise tests per test type
  results %>%
    mutate(
      t_p_adj = p.adjust(t_p, method = "BH"),
      w_p_adj = p.adjust(w_p, method = "BH"),
      pair = paste(group1, "vs", group2)
    )
}

# Summary: for each metric, the most significant pair
pairwise_summary <- function(pw_df) {
  pw_df %>%
    group_by(metric, metric_label) %>%
    slice_min(w_p, n = 1, with_ties = FALSE) %>%
    ungroup() %>%
    select(metric_label, pair, w_stat, w_p, w_p_adj, t_p) %>%
    mutate(
      sig = case_when(
        w_p_adj < 0.001 ~ "***",
        w_p_adj < 0.01  ~ "**",
        w_p_adj < 0.05  ~ "*",
        TRUE            ~ ""
      ),
      w_p = signif(w_p, 3),
      w_p_adj = signif(w_p_adj, 3),
      t_p = signif(t_p, 3),
      w_stat = round(w_stat, 1)
    ) %>%
    arrange(w_p) %>%
    kable(col.names = c("Metric", "Most Sig. Pair", "W stat", "Wilcox p",
                         "Wilcox p (BH)", "t-test p", "Sig"),
           caption = "Most significant pairwise comparison per metric (Wilcoxon)")
}

# Full pairwise table
pairwise_full_table <- function(pw_df) {
  pw_df %>%
    select(metric_label, pair, t_stat, t_p, t_p_adj, w_stat, w_p, w_p_adj) %>%
    mutate(
      across(c(t_stat, w_stat), ~round(.x, 2)),
      across(c(t_p, t_p_adj, w_p, w_p_adj), ~signif(.x, 3))
    ) %>%
    arrange(w_p) %>%
    kable(col.names = c("Metric", "Pair", "t stat", "t p", "t p (BH)",
                         "W stat", "W p", "W p (BH)"))
}
```

::: {.panel-tabset}

### 5-min

```{r pairwise-5min}
pw_5min <- run_pairwise_tests(metrics_5min, all_metric_cols)
```

**Summary — most significant pair per metric:**

```{r pairwise-5min-summary}
pairwise_summary(pw_5min)
```

<details>
<summary>Full pairwise results (click to expand)</summary>

```{r pairwise-5min-full}
pairwise_full_table(pw_5min)
```

</details>

### 1-min

```{r pairwise-1min}
pw_1min <- run_pairwise_tests(metrics_1min, all_metric_cols)
```

**Summary — most significant pair per metric:**

```{r pairwise-1min-summary}
pairwise_summary(pw_1min)
```

<details>
<summary>Full pairwise results (click to expand)</summary>

```{r pairwise-1min-full}
pairwise_full_table(pw_1min)
```

</details>

:::


## 6 PCA Analysis

```{r pca-helper}
run_pca_analysis <- function(df, metric_cols, title_prefix = "") {
  # Select metric columns and condition4
  df_pca <- df %>% select(id, condition4, all_of(metric_cols))

  # Report NA count
  n_na <- sum(!complete.cases(df_pca %>% select(all_of(metric_cols))))
  if (n_na > 0) {
    cat(sprintf("Note: %d subjects removed due to NA in PCA input columns.\n", n_na))
  }

  # Remove NAs for PCA
  df_complete <- df_pca %>% filter(complete.cases(across(all_of(metric_cols))))

  # Remove zero-variance columns
  df_num <- df_complete %>% select(all_of(metric_cols))
  zero_var <- names(df_num)[map_lgl(df_num, ~sd(.x, na.rm = TRUE) == 0)]
  if (length(zero_var) > 0) {
    cat(sprintf("Removed %d zero-variance columns: %s\n",
                length(zero_var), paste(zero_var, collapse = ", ")))
    metric_cols <- setdiff(metric_cols, zero_var)
    df_num <- df_complete %>% select(all_of(metric_cols))
  }

  stopifnot(nrow(df_num) >= 2, ncol(df_num) >= 2)

  # PCA
  pca_res <- prcomp(df_num, center = TRUE, scale. = TRUE)

  # Variance explained
  var_exp <- (pca_res$sdev^2) / sum(pca_res$sdev^2)
  var_pct <- round(var_exp * 100, 1)
  cum_var <- cumsum(var_pct)

  # Scores with metadata
  scores <- as.data.frame(pca_res$x) %>%
    bind_cols(df_complete %>% select(id, condition4))

  # --- Plot 1: PC1 vs PC2 ---
  p12 <- ggplot(scores, aes(x = PC1, y = PC2, color = condition4)) +
    geom_point(size = 2.5, alpha = 0.7) +
    scale_color_manual(values = condition4_colors) +
    labs(
      title = paste(title_prefix, "PC1 vs PC2"),
      x = sprintf("PC1 (%.1f%%)", var_pct[1]),
      y = sprintf("PC2 (%.1f%%)", var_pct[2]),
      color = "Condition"
    ) +
    theme(legend.position = "bottom")

  # --- Plot 2: PC3 vs PC4 ---
  p34 <- ggplot(scores, aes(x = PC3, y = PC4, color = condition4)) +
    geom_point(size = 2.5, alpha = 0.7) +
    scale_color_manual(values = condition4_colors) +
    labs(
      title = paste(title_prefix, "PC3 vs PC4"),
      x = sprintf("PC3 (%.1f%%)", var_pct[3]),
      y = sprintf("PC4 (%.1f%%)", var_pct[4]),
      color = "Condition"
    ) +
    theme(legend.position = "bottom")

  # --- Plot 3: Loadings heatmap (PC1-PC6) ---
  n_pcs <- min(6, ncol(pca_res$rotation))
  loadings_df <- as.data.frame(pca_res$rotation[, 1:n_pcs]) %>%
    rownames_to_column(var = "variable") %>%
    mutate(variable = str_remove(variable, "^avg_hr_")) %>%
    pivot_longer(cols = starts_with("PC"), names_to = "PC", values_to = "loading")

  pc_labels <- paste0("PC", 1:n_pcs, "\n(", var_pct[1:n_pcs], "%)")
  pc_lut <- setNames(pc_labels, paste0("PC", 1:n_pcs))

  p_load <- ggplot(loadings_df, aes(x = PC, y = variable, fill = loading)) +
    geom_tile() +
    scale_fill_gradient2(low = "cornflowerblue", mid = "white", high = "mediumvioletred",
                         midpoint = 0) +
    scale_x_discrete(labels = pc_lut) +
    labs(
      title = paste(title_prefix, "Loadings Heatmap"),
      x = "Principal Component", y = NULL, fill = "Loading"
    ) +
    theme(axis.text.y = element_text(size = 7))

  # --- Plot 4: Variance explained ---
  var_df <- tibble(
    PC = factor(paste0("PC", 1:min(15, length(var_pct))),
                levels = paste0("PC", 1:min(15, length(var_pct)))),
    pct = var_pct[1:min(15, length(var_pct))],
    cum = cum_var[1:min(15, length(var_pct))]
  )

  p_var <- ggplot(var_df, aes(x = PC, y = pct)) +
    geom_col(fill = "steelblue", alpha = 0.7) +
    geom_line(aes(y = cum, group = 1), color = "firebrick", linewidth = 1) +
    geom_point(aes(y = cum), color = "firebrick", size = 2) +
    geom_hline(yintercept = 80, linetype = "dashed", color = "grey50") +
    labs(
      title = paste(title_prefix, "Variance Explained"),
      x = NULL, y = "Variance (%)"
    ) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  list(p12 = p12, p34 = p34, p_load = p_load, p_var = p_var,
       pca_res = pca_res, scores = scores, n_na = n_na)
}
```

### 6.1 All Metrics (38)

::: {.panel-tabset}

#### 5-min

```{r pca-all-5min, fig.width=10, fig.height=6}
pca_all_5 <- run_pca_analysis(metrics_5min, all_metric_cols, "All Metrics —")
pca_all_5$p12
```

```{r pca-all-5min-34, fig.width=10, fig.height=6}
pca_all_5$p34
```

```{r pca-all-5min-load, fig.width=10, fig.height=10}
pca_all_5$p_load
```

```{r pca-all-5min-var, fig.width=8, fig.height=4}
pca_all_5$p_var
```

#### 1-min

```{r pca-all-1min, fig.width=10, fig.height=6}
pca_all_1 <- run_pca_analysis(metrics_1min, all_metric_cols, "All Metrics —")
pca_all_1$p12
```

```{r pca-all-1min-34, fig.width=10, fig.height=6}
pca_all_1$p34
```

```{r pca-all-1min-load, fig.width=10, fig.height=10}
pca_all_1$p_load
```

```{r pca-all-1min-var, fig.width=8, fig.height=4}
pca_all_1$p_var
```

:::

### 6.2 avg_hr Metrics Only (34)

::: {.panel-tabset}

#### 5-min

```{r pca-avghr-5min, fig.width=10, fig.height=6}
pca_avghr_5 <- run_pca_analysis(metrics_5min, avg_hr_cols, "avg_hr —")
pca_avghr_5$p12
```

```{r pca-avghr-5min-34, fig.width=10, fig.height=6}
pca_avghr_5$p34
```

```{r pca-avghr-5min-load, fig.width=10, fig.height=10}
pca_avghr_5$p_load
```

```{r pca-avghr-5min-var, fig.width=8, fig.height=4}
pca_avghr_5$p_var
```

#### 1-min

```{r pca-avghr-1min, fig.width=10, fig.height=6}
pca_avghr_1 <- run_pca_analysis(metrics_1min, avg_hr_cols, "avg_hr —")
pca_avghr_1$p12
```

```{r pca-avghr-1min-34, fig.width=10, fig.height=6}
pca_avghr_1$p34
```

```{r pca-avghr-1min-load, fig.width=10, fig.height=10}
pca_avghr_1$p_load
```

```{r pca-avghr-1min-var, fig.width=8, fig.height=4}
pca_avghr_1$p_var
```

:::

### 6.3 Top Significant avg_hr Metrics

```{r select-top-metrics}
# Select avg_hr metrics where KW p_adj < 0.05 (from 5-min omnibus)
top_metrics_kw <- omnibus_5min %>%
  filter(test == "Kruskal-Wallis", metric %in% avg_hr_cols) %>%
  filter(p_adj < 0.05) %>%
  pull(metric)

# Fallback: if fewer than 3 pass threshold, use top 10 by raw p-value
if (length(top_metrics_kw) < 3) {
  cat("Fewer than 3 metrics pass KW p_adj < 0.05 — using top 10 by raw p-value.\n")
  top_metrics_kw <- omnibus_5min %>%
    filter(test == "Kruskal-Wallis", metric %in% avg_hr_cols) %>%
    arrange(p.value) %>%
    slice_head(n = 10) %>%
    pull(metric)
}

cat(sprintf("Using %d significant avg_hr metrics for PCA:\n", length(top_metrics_kw)))
cat(paste("-", str_remove(top_metrics_kw, "^avg_hr_")), sep = "\n")
```

::: {.panel-tabset}

#### 5-min

```{r pca-top-5min, fig.width=10, fig.height=6}
pca_top_5 <- run_pca_analysis(metrics_5min, top_metrics_kw, "Top Sig. —")
pca_top_5$p12
```

```{r pca-top-5min-34, fig.width=10, fig.height=6}
pca_top_5$p34
```

```{r pca-top-5min-load, fig.width=10, fig.height=8}
pca_top_5$p_load
```

```{r pca-top-5min-var, fig.width=8, fig.height=4}
pca_top_5$p_var
```

#### 1-min

```{r pca-top-1min, fig.width=10, fig.height=6}
pca_top_1 <- run_pca_analysis(metrics_1min, top_metrics_kw, "Top Sig. —")
pca_top_1$p12
```

```{r pca-top-1min-34, fig.width=10, fig.height=6}
pca_top_1$p34
```

```{r pca-top-1min-load, fig.width=10, fig.height=8}
pca_top_1$p_load
```

```{r pca-top-1min-var, fig.width=8, fig.height=4}
pca_top_1$p_var
```

:::


## 7 Summary

```{r summary-text, results='asis'}
# Top 5 metrics by omnibus KW (5-min)
top5_kw <- omnibus_5min %>%
  filter(test == "Kruskal-Wallis") %>%
  arrange(p_adj) %>%
  slice_head(n = 5)

cat("### Key Findings\n\n")
cat(sprintf("**Omnibus tests**: %d of %d metrics show significant group differences (KW p_adj < 0.05, 5-min resolution).\n\n",
            sum(omnibus_5min$test == "Kruskal-Wallis" & omnibus_5min$p_adj < 0.05, na.rm = TRUE),
            length(all_metric_cols)))

cat("**Top 5 most discriminating metrics** (Kruskal-Wallis, 5-min, BH-adjusted):\n\n")
for (i in seq_len(nrow(top5_kw))) {
  cat(sprintf("%d. **%s** — p_adj = %.2e %s\n",
              i, top5_kw$metric_label[i], top5_kw$p_adj[i], top5_kw$sig[i]))
}

# Pairwise highlight
top_pair <- pw_5min %>%
  slice_min(w_p, n = 1, with_ties = FALSE)
cat(sprintf("\n**Most significant pairwise comparison**: %s vs %s on metric %s (Wilcoxon p = %.2e).\n\n",
            top_pair$group1, top_pair$group2,
            str_remove(top_pair$metric, "^avg_hr_"),
            top_pair$w_p))

cat("**Note**: CAD_Diabetes (n=29) and ESRD (n=41) groups have relatively small sample sizes; ",
    "non-parametric tests (Kruskal-Wallis, Wilcoxon) are more reliable than parametric alternatives ",
    "for these groups.\n")
```
